---
title: "Rental Vacancy Charting"
author: "Alec A. Nathan"
output: html_document
---

###Setup

  Load your required packages and API keys. As a practice, please store API keys in your .Renviron this will prevent you from accidentally uploading sensitive information to a code repository.
  
```{r setup, include=FALSE}
library(tidyverse)
library(tidyquant)
library(timetk)
library(plotly)
library(janitor)
library(ggplot2)

#Macro Factor Libraries
library(fredr)
library(tidycensus)

#Data Visualization Extras
library(png)
library(ggthemes)
library(showtext)


#Load Passwords and Keys from Environment
FRED_API_KEY <- 
  Sys.getenv("FRED_API_KEY")

CENSUS_API_KEY <- 
  Sys.getenv("CENSUS_API_KEY")

#Set Period For Observations
start_period <- ymd('2000-01-01')


knitr::opts_chunk$set(echo = TRUE)
```

###Load Data from FRED

Load in our raw data from FRED. If you rerun this script in three months another quarter of data will be appended.

```{r}
vacancy_data_fred_tbl <- fredr_series_observations(series_id = "RRVRUSQ156N") %>% 
  select(date,value) %>% 
  filter(date >= start_period) %>% 
  #Values come in as whole numbers, not percentages. We must mutate our value column to the appropriate scale.
  mutate(value = value/100)
```

###Simple Chart

If we define our axes and pass through a geometry we can create a very simple chart.

```{r}
simple_chart <- vacancy_data_fred_tbl %>% 
  ggplot(aes(x=date,y=value))+
  geom_line()
```

###Add Axis Formatting & Titles

Names and formatting help users interpret data

```{r}
formatted_axes <- simple_chart +
  scale_x_date(name         = "DATE",
               date_breaks  = "3 years",
               date_labels  = "%Y",
               minor_breaks = NULL)+
  scale_y_continuous(name   = "VACANCY PERCENT",
                     labels = scales::percent)
```

Adding a title allows users to contextualize the data and ascertain the source of the data.

```{r}
added_title <- formatted_axes +
  ggtitle(label    = "RENTAL VACANCY RATE IN THE UNITED STATES",
          subtitle = "RETRIEVED FROM FRED, FEDERAL RESERVE BANK OF ST. LOUIS")
```

###Example themes and applying the Minimal Theme

Prebuilt themes in GGPLOT2 are a great start, but not the end-all, be-all of a bespoke reporting solution.

```{r}
ggthemes_example_538 <- added_title + 
  theme_fivethirtyeight()

ggthemes_example_economist <- added_title +
  theme_economist()

ggplot_minimal <- added_title +
  theme_minimal()
```


###Recession Shading Function

The function below can be added as a layer to a gg object. Thank you Jonathan Regenstein!

```{r}
#Load Recession Shading for Graphs
recession_shade_fun <- function(color_chosen = "darkgray", start_date = start_period){ 
  "USREC" %>% 
  tq_get(get = "economic.data", from = start_date) %>% 
  select(date, recession_flag = price) %>% 
  mutate(recession_label = case_when(recession_flag == 1 & lag(recession_flag == 0) ~ str_glue("{year(date)} recession"), 
         TRUE ~ NA_character_)) %>% 
  filter(recession_flag == 1) %>% 
  fill(recession_label, .direction = "down") %>% 
  group_by(recession_label) %>% 
  slice(1, n()) %>% 
  select(-recession_flag) %>% 
  mutate(start = case_when(date == min(date) ~ date,
                           TRUE ~ NA_Date_),
         end = case_when(date == max(date) ~ date,
                         TRUE ~ NA_Date_),
         end = lead(end)
         ) %>% 
  filter(!is.na(start)) %>% 
  select(-date) %>% 
    geom_rect(data = ., 
            inherit.aes = F, 
            aes(xmin = start, 
                xmax = end, 
                ymin = -Inf, 
                ymax = +Inf), 
            fill = color_chosen, 
            alpha = 0.4)
}
```

###Adding More Formatting

Let's add formatting to our plot. The theme function gives you endless customization capabilities.

```{r}
ggplot_w_correct_line <- vacancy_data_fred_tbl %>% 
  ggplot(aes(x=date,y=value)) +
  geom_line(color     = "#9BC2E6",
            linewidth = 2) +
  scale_x_date(name         = "DATE",
               date_breaks  = "3 years",
               date_labels  = "%Y",
               minor_breaks = NULL)+
  scale_y_continuous(name   = "VACANCY PERCENT",
                     labels = scales::percent) +
  ggtitle(label    = "RENTAL VACANCY RATE IN THE UNITED STATES",
          subtitle = "RETRIEVED FROM FRED, FEDERAL RESERVE BANK OF ST. LOUIS")+
  theme_minimal()+
  theme(panel.grid       = element_line(colour      = "white"),
        panel.background = element_rect(fill        = "white"),
        panel.border     = element_rect(colour      = "white",
                                    fill            = NA),
        plot.background  = element_rect(fill        = "#E8F2FA",
                                       colour       = "#E8F2FA"))+
  recession_shade_fun()+
  labs(caption = "RECESSIONS SHADED IN GREY")
```

###Loading Fonts to Use

Let's prepare and load our Google Font.

```{r}
trace(grDevices::png, exit = quote({
    showtext::showtext_begin()
}), print = FALSE)

font_add_google(name = "Montserrat", family = "Montserrat")
```

###Applying Fonts

Let's pass our newly loaded font through our theme. The font family is the most important attribute here.

```{r}
ggplot_w_font <- ggplot_w_correct_line+
  theme(text = element_text(family = "Montserrat"))
```

###Add Logo to Local Environment

Store your logo in your directory. From here the get_png function will make it accessible to GGPlot2.

```{r}
get_png <- function(filename) {
  grid::rasterGrob(png::readPNG(filename), interpolate = TRUE)
}

logo <- get_png("linkedinaan.png")
test_logo <- grid::roundrectGrob()
```

###Place Logo and Relocate Caption

Our caption and logo are overlapping. Le'ts use the plot.caption and plot.caption.position arguments to move the caption to the left.

```{r}
ggplot_final <- ggplot_w_font +
  annotation_custom(logo,
                    xmin = ymd('2016-07-01'),
                    xmax = ymd('2022-12-31'),
                    ymin = .025,
                    ymax = .05)+
  coord_cartesian(clip = "off")+
  theme(plot.margin            = unit(c(1,1,2.75,1),"lines"),
        plot.caption.position = "plot",
        plot.caption          = element_text(hjust = 0))
```
